---
title: "Novartis Data Science Hiring challenge"
output: 
    html_document
---  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, warning=FALSE,message=F}
{
library(tidyverse)
library(lubridate)
library(plotly)
library(caret)
library(randomForest)
library(h2o)
library(highcharter)
library(hrbrthemes)
library(viridis)
library(purrr)
library(gridExtra)
library(formatR)
library(tidytext)
library(RColorBrewer)
library(DT)  
library(Amelia)
library(corrplot) 
library(missRanger)
}
```

### Import the train and test dataset
```{r}
train<-read.csv("Train.csv",header = T, na.strings = c(""," ","NA"))
test<-read.csv("Test.csv",header = T, na.strings = c(""," ","NA"))
```

### Structure the Train data
```{r}
str(train)
```

### Summary of Train data
```{r}
summary(train)
```

### Structure of Test data
```{r}
str(test)
```


### Summary of Test data
```{r}
summary(test)
```

***
#### To make a exploratory data analysis task more informative, recode the target variable(MULTIPLE_OFFENSE) levels "1" as Yes and "0" as No.

```{r}

train$MULTIPLE_OFFENSE <- ifelse(train$MULTIPLE_OFFENSE == 1, "Yes", "No")

```


*** 
#### The following plot shows Number hack attacks on digital payments from the year 1991 to 2018. 

```{r , error=FALSE,warning=FALSE}
train %>% group_by(MULTIPLE_OFFENSE) %>% summarise(counts = n()) %>%  plot_ly(
  labels = ~ MULTIPLE_OFFENSE,
  values = ~ counts,
  type = 'pie'
  ) %>% layout(legend = list(
  orientation = "h",
  xanchor = "center",
  x = 0.5
  ))
  
```

***

#### Let visualize the train & test data columns from X_1 to X_15 and see the data distribution of them 



### {.tabset}



#### Train data {-}



```{r, warning = FALSE, message = F,fig.width=7, fig.height=10}

ggplotly(
  train %>% select(-MULTIPLE_OFFENSE) %>%
  gather(measurement, value, X_1:X_15, factor_key = T) %>%
  ggplot(aes(x = value, fill = measurement)) + geom_histogram()  +
  facet_wrap(~ measurement , scales = "free" , ncol = 3) + theme_light() +
  theme(
  legend.position = "none",
  panel.spacing = unit(2, "lines"),
  axis.title.x = element_blank(),
  axis.title.y = element_blank()
  )
  )
```



#### Test data {-}



```{r,warning=FALSE,message=F,fig.width=10, fig.height=10}
ggplotly(
  test %>%
  gather(measurement, value, X_1:X_15, factor_key = T) %>%
  ggplot(aes(x = value, fill = measurement)) + geom_histogram()  +
  facet_wrap(~ measurement, scales = "free", ncol = 3) + theme_light() +
  theme(
  legend.position = "none",
  panel.spacing = unit(2, "lines"),
  axis.title.x = element_blank(),
  axis.title.y = element_blank()
  )
  )
```



### {-}


****






#### Train and Test data are having date column so here to perform some feature engineering task to extract the valuable information from the date column.Before that to combine the train and test data and stored it in a new dataframe.

```{r}
fulldata <- merge(train, test, all = T)
```


#### The format of date in dataset is character so convert it to date type


```{r}
fulldata$DATE  <-dmy(fulldata$DATE)
fulldata<-data.frame(fulldata)
```


#### Next extract the features like day, month, year, week ,weekday,quarter and leap year from date column


```{r , warning=FALSE,message=F}
{
  fulldata$date_day <- lubridate::day(fulldata$DATE)#extract day
  fulldata <- fulldata %>%  select(1:2, 19, everything())#column reorder
  
  fulldata$date_month <- lubridate::month(fulldata$DATE, label=T)#extract month
  fulldata <- fulldata %>%  select(1:3, 20, everything())#column reorder
  
  fulldata$date_year <- lubridate::year(fulldata$DATE)#extract year
  fulldata <- fulldata %>%  select(1:4, 21, everything())#column reorder
  
  fulldata <- fulldata %>% group_by(DATE) %>% mutate(date_count = n())#Date count
  fulldata <- fulldata %>%  select(1:2, 22, everything())#column reorder
  
  
  fulldata$date_quarter <- lubridate::quarter(fulldata$DATE)#extract quarter
  fulldata <- fulldata %>%  select(1:6, 23, everything())#column reorder
  
  fulldata$date_week <- lubridate::week(fulldata$DATE)#extract week
  fulldata <- fulldata %>%  select(1:7, 24, everything())#column reorder
  
  fulldata$date_wday <- lubridate::wday(fulldata$DATE , label=T )#extract weekday
  fulldata <- fulldata %>%  select(1:8, 25, everything())#column reorder
  
  
  fulldata$date_leap_year <- lubridate::leap_year(fulldata$DATE)#extract leap year info
  fulldata <- fulldata %>%  select(1:9, 26, everything())#column reorder
}
```

```{r}
datatable(head(fulldata[,c(2:10)]), class = 'cell-border stripe')
```


#### The following time series chart explains how the number of hack attacks on digital payments changed over the time.


```{r,warning=FALSE,message=F}
#filtering train data from fulldata
highchart() %>%
  hc_add_series(
  data =
  fulldata %>% filter(!is.na(MULTIPLE_OFFENSE)) %>%
  group_by(date_year, MULTIPLE_OFFENSE) %>%
  summarise(counts = n()) ,
  type = "line",
  hcaes(x = date_year,
  y = counts,
  group = MULTIPLE_OFFENSE)
  )
```



#### lets break down the time series plot into various levels and see how the number of hack attacks on digital payments changed in  weeks, months, quarters and leap years from the year 1991 to 2018.

#### The following visualization explains the number hack attacks in each month over the year 

```{r,warning=FALSE,message=F,fig.width=10, fig.height=10}
ggplotly(
  fulldata %>% drop_na(MULTIPLE_OFFENSE) %>%
  filter(MULTIPLE_OFFENSE == 'Yes') %>%
  group_by(date_month, date_year, MULTIPLE_OFFENSE) %>%
  summarise(counts = n())  %>%
  ggplot(
  aes(
  x = reorder_within(date_month,-counts,
  date_year),
  y = counts,
  fill = date_month,
  text = paste("Month:",
  date_month,
  "<br> counts:",
  counts)
  )
  ) +
  geom_bar(stat = "identity",
  position = "stack") + scale_fill_brewer(palette = "Set3") + xlab("Month") +
  facet_wrap( ~ date_year , scales = "free" , ncol = 3) +
  scale_x_reordered() +
  theme(
  panel.spacing = unit(1, "lines"),
  axis.text.x = element_blank(),
  axis.title.x = element_text(face = "bold", hjust = 0.5),
  axis.title.y = element_blank()
  ) ,
  tooltip = "text"
  ) %>%
  layout(legend = list(
  orientation = "h",
  x = 0,
  y = -0.05
  ))

```

#### The following visualization explains the number hack attacks in each quarter over the year.In the 2nd & 3rd quarter, more number of hack attacks occurred

```{r,warning=FALSE,message=F,fig.width=10, fig.height=10}
ggplotly(
  fulldata %>% drop_na(MULTIPLE_OFFENSE) %>%
  filter(MULTIPLE_OFFENSE == 'Yes') %>%
  dplyr::group_by(date_quarter, date_year, MULTIPLE_OFFENSE) %>%
  summarise(counts = n())  %>%
  ggplot(aes(
  x = reorder_within(date_quarter,-counts,
  date_year),
  y = counts,
  fill = factor(date_quarter),
  text = paste("Quarter:",
  date_quarter,
  "<br> counts:",
  counts)
  )) +
  geom_bar(stat = "identity",
  position = "stack") + scale_fill_brewer(palette = "Accent") + xlab("Quarter") +
  facet_wrap( ~ date_year , scales = "free" , ncol = 3) +
  scale_x_reordered() +
  theme(
  panel.spacing = unit(1, "lines"),
  axis.text.x = element_text(fac = "bold"),
  axis.title.x = element_text(face = "bold", hjust = 0.5),
  axis.title.y = element_blank()
  ),
  tooltip = "text"
  )  %>%
  layout(legend = list(
  orientation = "h",
  x = 0.3,
  y = -0.05
  ))  

```


#### The following plot explains number hack attacks in each day of the week over the year, the most hack attacks oocured in week days.

```{r,warning=FALSE,message=F,fig.width=10, fig.height=10}
ggplotly(
  fulldata %>% drop_na(MULTIPLE_OFFENSE) %>%
  filter(MULTIPLE_OFFENSE == 'Yes') %>%
  group_by(date_wday, date_year, MULTIPLE_OFFENSE) %>%
  summarise(counts = n())  %>%
  ggplot(aes(
  x = reorder_within(date_wday,-counts,
  date_year),
  y = counts,
  fill = date_wday,
  text = paste("Week Day:",
  date_wday,
  "<br> counts:",
  counts)
  )) +
  geom_bar(stat = "identity",
  position = "stack") + scale_fill_brewer(palette = "Set2") + xlab("Week Day") +
  facet_wrap( ~ date_year , scales = "free" , ncol = 3) +
  scale_x_reordered() +
  theme(
  panel.spacing = unit(1, "lines"),
  axis.text.x = element_blank(),
  axis.title.x = element_text(face = "bold", hjust = 0.5),
  axis.title.y = element_blank()
  ),
  tooltip = "text"
  ) %>%
  layout(legend = list(
  orientation = "h",
  x = 0.2,
  y = -0.05
  ))
```


#### Ordinary Year - Leap Year hack attacks 

```{r,warning=FALSE,message=F ,fig.width=9, fig.height=10}
fulldata %>% drop_na(MULTIPLE_OFFENSE) %>%
  filter(MULTIPLE_OFFENSE == 'Yes') %>%
  group_by(date_leap_year, date_year) %>%
  summarise(counts = n())  %>%
  ggplot(aes(
  x = date_year,
  y = counts,
  label = date_year,
  color = date_leap_year
  )) +
  geom_point() +
  geom_text(aes(color = factor(date_leap_year))) +
  geom_line(aes(
  x = date_year,
  y = counts,
  color = date_leap_year,
  group = 1
  ),
  size = 1) +
  xlab("Year") +
  scale_x_continuous(breaks = seq(1991, 2018, by = 1)) +
  labs(color = "Leap Year\n") +
  theme(
  legend.position = "bottom",
  axis.text.x = element_blank(),
  axis.title.y = element_blank()
  )
```
 


#### The X_1 to X_15 features are anonymous so not able to identify which one is quantitative and qualitative feature.Lets assume the features X_1 to X_15 as continuous type and see the correlation of those features.



### {.tabset}


#### Train data {-}



```{r,warning=F,error=FALSE}
cor<-cor( train %>% na.omit() %>%  select_if(is.numeric ))
corrplot(cor, method = "circle")
```



#### Test data {-}



```{r,warning=F,error=FALSE}
cor1<-cor( test %>% na.omit() %>%  select_if(is.numeric ))
corrplot(cor1, method = "circle")
```


### {-}

##### In the above corrrelation plot blue color represents positive correlation and red color represents negative correlation ,the color intensity explains how much the features are correlated from -1 to +1.(0 means no correlation,1 means positive correlation, -1 means negative correlation).
##### There is a strong positive correlation between X_2 and X_3 , X_10 and X_12.




#### lets visualize the Missing values from train and test data 




***


### {.tabset}



#### Train data {-}



```{r}
missmap(train[,-c(1,2,18)])
```



#### Test data {-}



```{r}
missmap(test[,-c(1,2)])
```



### {-}


***




#### Using missRanger package to impute missing values.

```{r ,warning=F,error=FALSE}


fulldataimp<-missRanger(fulldata[,-c(1,2,26)], num.trees = 100,verbose = T,
                        seed = 887,pmm.k = 4)

```


#### insert incident id to new data frame



```{r ,warning=F,error=FALSE}

fulldataimp$INCIDENT_ID<-fulldata$INCIDENT_ID

fulldataimp<-fulldataimp %>% select(24,everything())

```

#### insert MULTIPLE_OFFENSE to new data frame

```{r ,warning=F,error=FALSE}

fulldataimp$MULTIPLE_OFFENSE<-fulldata$MULTIPLE_OFFENSE

```


#### Lets recode the MULTIPLE_OFFENSE by 1 and 0

```{r ,warning=F,error=FALSE}

fulldataimp$MULTIPLE_OFFENSE<-as.factor(ifelse(fulldataimp$MULTIPLE_OFFENSE=='Yes',1,0))

```

#### lets split the data into  Train set, validation set, Test set 

#### Train set


```{r ,warning=F,error=FALSE}

trainn<-fulldataimp[complete.cases(fulldataimp$MULTIPLE_OFFENSE),]

```

#### Test set

```{r ,warning=F,error=FALSE}

testn<-fulldataimp[!complete.cases(fulldataimp$MULTIPLE_OFFENSE),]

```

#### Create a stratified sample train and validation set data from trainn dataframe

```{r ,warning=F,error=FALSE}

index<-createDataPartition(y=trainn$MULTIPLE_OFFENSE ,p=0.75,list = F)

trains<-trainn[index,]

vals<-trainn[-index,]

```

#### Classify hack attacks by using Random forest algorithm , removed high correlated feature X_2 and X_10

```{r ,warning=F,error=FALSE}

rfmod<-randomForest(factor(MULTIPLE_OFFENSE)~., data=trains[,c(10,11,13:20,22:25)], strata=MULTIPLE_OFFENSE,
                importance=T,ntree=500,mtry=4,replace=F)


```

#### Evaluate the model performance by using Validation dataset

```{r ,warning=F,error=FALSE}

valpre<-predict(rfmod,vals)

```

#### create a confusion matrix and see how its predicted the actual classes

```{r ,warning=F,error=FALSE}

cm<-confusionMatrix(valpre, factor(vals$MULTIPLE_OFFENSE),mode="everything") 

```

#### Validation set accuaracy

```{r ,warning=F,error=FALSE}

cm$byClass[7] 

```
*****